---
title: "Single Cell QC"
author: "Dylan Rausch"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
params:
  par: "default value"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(tidyverse.quiet = TRUE)
```

This report contains the results of a minimal `targets` data analysis pipeline. We reference our targets with `tar_load()` and `tar_read()`. This ensures

1. Because of the `tar_render()` function from the [`tarchetypes`](https://wlandau.github.io/tarchetypes) package (see `_targets.R`) `targets` automatically detects the dependencies of this report and rebuilds it when its dependencies change.
1. We can run the report by itself if the targets are already in the `_targets/` data store.

Here is our fitted regression model of ozone versus wind and temperature.

```{r}
library(targets)
```

Load the SCE object of interest
```{r}
# parameter is defined in the YAML header (specifies branch object to select)
branch <- params$par
tar_load(sce_objects, branches=branch)
sce_objects <- sce_objects[[1]]
sce_objects
```

```{r}
# Global classification results
table(sce_objects$hto_global)
```

```{r}
table(sce_objects$hto_global, sce_objects$discard)
```

# Visualize enrichment for selected HTOs with ridge plots (rebuild seurat object with HTO)
```{r warning=FALSE, fig.width=10, fig.height=5}
seur.hto <- as.Seurat(altExp(sce_objects, "hto"), counts = "counts", data=NULL)
seur.hto <- NormalizeData(seur.hto, normalization.method = "CLR")
seur.hto$hash_id <- sce_objects$hash_id
seur.hto$hto_global <- sce_objects$hto_global
seur.hto$hash_max <- sce_objects$hash_max

# Group cells based on the max HTO signal
Idents(seur.hto) <- "hash_max"
RidgePlot(seur.hto, features = rownames(seur.hto), ncol = 3)
```

```{r include=FALSE}
plot <- determine_quantile(seur.hto)
```

# Analyze quantile selection
```{r}
plot[[1]] %>% pivot_longer(-quant) %>% filter(quant!=1) %>% ggplot() + 
    aes(x=quant, y=value, color=name) + geom_point(size=0.5) + geom_line() +
    geom_vline(xintercept = plot[[2]], linetype="dashed")
```

# Analyze intra-sample doublets
```{r}
# calculate number of neighbors at each proportion that are doublets
sce_objects$proportion_dub_neighbors %>% 
    data.frame("prop"=.) %>% 
    group_by(prop) %>% 
    summarize(n=n()) %>% 
    mutate(pct = n/sum(n)) -> data 

# find point at which we gain very few doublets as proportion increases 
cut <- data$prop[PCAtools::findElbowPoint(variance = sort(data$n, decreasing = T))+1]

# plot number of neighbors at different proportions
ggplot(data) + 
  aes(x=prop, y=n) +
  geom_point() +
  geom_line() +
  geom_vline(xintercept = cut)
```

# look at which cells are filtered from which groups
```{r fig.width=10, fig.height=5}
gridExtra::grid.arrange(
    plotColData(sce_objects, x="hash_max", y="sum", colour_by="discard") +
        scale_y_log10() + ggtitle("Total count"),
    plotColData(sce_objects, x="hash_max", y="detected", colour_by="discard") +
        scale_y_log10() + ggtitle("Detected features"),
    plotColData(sce_objects, x="hash_max", y="sum", colour_by="hto_global") + 
      scale_y_log10() + ggtitle("Hash Calls"),
    plotColData(sce_objects, x="hto_global", y="sum", colour_by = "hto_global") + 
      scale_y_log10() + ggtitle("Total Count"),
    ncol=2
)
```

# Visualize distribution of counts across hashes
```{r}
plotColData(sce_objects, x="hto_global", y="detected", colour_by="discard", other_fields = "hash_max") +
        scale_y_log10() + ggtitle("Total count") +
  facet_wrap(~hash_max)
```

# Identify location of doublets in TSNE space, also location of intra-sample doublets
```{r fig.width=10, fig.height=4}
sce_objects <- runTSNE(sce_objects, dimred="PCA")
gridExtra::grid.arrange(
    plotTSNE(sce_objects, colour_by="proportion_dub_neighbors") + 
      ggtitle("Doublet proportions") +
      theme(legend.position = "none"),
    plotTSNE(sce_objects, colour_by="doublet") + ggtitle("Known doublets"),
    ggcells(sce_objects) +
        geom_point(aes(x=TSNE.1, y=TSNE.2), color="grey") +
        geom_point(aes(x=TSNE.1, y=TSNE.2), color="red", 
            data=function(x) x[x$predicted_dub_cut,]) +
        ggtitle("Predicted intra-sample doublets") +
      theme(legend.position = "none"),
    ncol = 3     
)
```

# Generate a two dimensional tSNE embedding for HTOs.Here we are grouping cells by singlets and doublets for simplicity.
```{r}
Idents(seur.hto) <- "hto_global"
# Calculate a distance matrix using HTO
hto.dist.mtx <- as.matrix(dist(t(GetAssayData(object = seur.hto))))
# Calculate tSNE embeddings with a distance matrix
seur.hto <- RunTSNE(seur.hto, distance.matrix = hto.dist.mtx, perplexity = 100)
DimPlot(seur.hto)
```




